\section{Discussion and conclusion}

The results show that 2D oriented-powder WAXS patterns can be analyzed quantitatively without reducing the image to either a single-crystal indexing problem or a conventional 3D powder pattern. The forward model keeps detector geometry, sample alignment, mosaic broadening, ordered structure, and stacking disorder within one detector-space calculation, while the refinement separates these terms by the observables that constrain them most directly. Ordered Bi$_2$Te$_3$ and Bi$_2$Se$_3$ provide the baseline case where, once geometry and mosaicity are fixed, the remaining Bragg-intensity variation can be assigned to ordered-structure parameters. PbI$_2$ provides the disorder case. The ordered Bragg model reproduces the peak positions and primary Bragg intensities but leaves structured rod intensity that is accounted for by adding the stacking-fault correction.

The fitted structure-factor parameters provide an additional consistency check on this separation of model terms. For Bi$_2$Te$_3$ and Bi$_2$Se$_3$, the final detector-space comparisons in Fig.~\ref{fig:ordered_detector_fits} were obtained with lattice constants, internal coordinates, occupancies, and directional displacement terms that remain close to the cited structural ranges summarized in Table~\ref{tab:ordered_sf_parameters}. The Bi$_2$Te$_3$ fit is consistent with the relaxed MBE lattice scale and the bulk-neutron coordinate and displacement-parameter convention, while the Bi$_2$Se$_3$ fit remains compatible with the reported anisotropic-model lattice, Bi/Se coordinates, outer-Se occupancy, and constrained Se displacement terms.\cite{Mansour2014,Steiner2014,Gharsallah2016,Callaert2019,Trueblood1996} This agreement is not a proof of uniqueness, because detector-space intensities remain correlated with scale, profile tails, absorption, and the selected reflection set. It does show that the ordered-film simulations do not require unphysical changes in the average structures to reproduce the measured intensity patterns.

The PbI$_2$ refinement extends the same logic to a film where the ordered average structure is not sufficient. The ordered PbI$_2$ baseline in Fig.~\ref{fig:pbi2_disorder_fit} uses 2H lattice constants, the corrected iodine coordinate, and near-full occupancies consistent with the stoichiometric average model in Table~\ref{tab:pbi2_sf_parameters}.\cite{Minagawa1975,Palosz1990,Lin2019} The fitted Pb and I displacement terms are best interpreted as sample-specific ordered-baseline parameters rather than transferable PbI$_2$ constants, because site-resolved anisotropic displacement values for PbI$_2$ are not widely or consistently reported in a directly interchangeable convention. Within that convention-sensitive comparison, the fitted values preserve the qualitative anisotropy expected for the layered structure and prevent the stacking-fault term from compensating for simple ordered-intensity errors.\cite{Minagawa1975,Trueblood1996}

Peak positions determine geometry before peak widths are interpreted. Local profiles determine the mosaic and instrumental response before integrated intensities are assigned to structure. Selected $Q_r/Q_z$ rod profiles determine stacking disorder only after the ordered baseline is fixed. That order reduces the risk that diffuse scattering is incorrectly absorbed into geometry, mosaicity, displacement parameters, or independent peak areas. In the PbI$_2$ case, the fitted disorder values, $p=0.18$, $N=32$ layers, and a 2H:6H stacking-sequence fraction of $0.76:0.24$, should therefore be read as effective descriptors of the selected rod and fitted stacking model rather than universal PbI$_2$ constants.

The approach is also limited by the information retained in the selected observables. A detector-space forward model cannot remove ambiguity if the chosen reflections do not span enough geometry, mosaic, and intensity constraints. The stacking-disorder parameters are therefore listed as effective descriptors of the selected rod family and fitted model class. Within that scope, the combination of detector-space fitting, physically bounded ordered-structure parameters, and selected-rod validation provides a practical route for comparing ordered and faulted layered films using area-detector WAXS data.
