\section{Refinement workflow}

With the forward model defined above, refinement proceeds sequentially from instrumental quantities to sample-specific scattering parameters in order to reduce correlations. Beam phase-space terms are determined first, followed by detector calibration from a dark-subtracted hBN standard. The sample-side analysis then uses dark-subtracted WAXS images from the materials of interest: ordered-film images for Bi$_2$Se$_3$ and Bi$_2$Te$_3$, and diffuse-scattering images for PbI$_2$. Positional information dominates the alignment stage, where indexed Bragg peaks are matched by least squares, whereas line-shape and intensity information dominate the later stages. Mosaicity is obtained from a joint fit to two images acquired at different $\theta_i$, after which ordered-structure and disorder parameters are refined with geometry and mosaicity held fixed. The staged protocol is summarized in Table~\ref{tab:param_effects}; additional implementation detail is provided in the supporting information (Sec.~S1).

\begin{table}[!ht]
  \centering
  \caption{Staged refinement workflow. For each step we list the data used, the parameters fit, and the dominant observables that constrain them from the area-detector measurements.}
  \label{tab:param_effects}
  \scriptsize
  \setlength{\tabcolsep}{2.5pt}
  \renewcommand{\arraystretch}{1.05}
  \begin{tabularx}{\linewidth}{@{}c l L L L@{}}
    \toprule
    Step & Stage & Data used & Parameters fit & Primary observables \\
    \midrule
    1 &
    Beam characterization (direct beam) &
    Direct-beam images at several detector distances &
    Beam widths $w_{x,z}$ and divergences $\Delta\theta_{x,z}$ &
    Spot centroid and width versus detector distance (instrument broadening). \\
    2 &
    Detector geometry calibration (3D powder standard) &
    Dark-subtracted hBN calibrant image &
    Detector distance $D$, tilts $(\beta,\gamma)$, in-plane rotation $\chi$, beam center $(x_0,y_0)$ &
    Ring center, ellipticity, and orientation (pixel-to-$\mathbf{Q}$ mapping). \\
    3 &
    Sample alignment and goniometer misalignment &
    Dark-subtracted sample image &
    Sample geometry $(\theta_i,\delta,z_B,z_S)$, goniometer misalignment $(\alpha,\psi)$, film $(W,H,t)$ &
    Least-squares residuals of indexed specular and off-specular Bragg-peak positions. \\
    4 &
    Mosaicity from Bragg-peak shapes (two $\theta_i$ images) &
    Two dark-subtracted images of the same sample at different $\theta_i$ &
    Mosaic kernel $(\eta,\Gamma,\sigma)$ &
    Arc/cap widths and specular off-condition tails within local reflection ROIs. \\
    5 &
    Ordered-structure refinement &
    Dark-subtracted Bi$_2$Se$_3$/Bi$_2$Te$_3$ images &
    Scale and structure-factor parameters ($o_a$, $\mathbf{r}_a$, $\Delta\mathbf{r}_a$, $\mathbf{U}$) &
    Relative Bragg intensities and rod profiles across the full 2D images. \\
    6 &
    Disorder refinement &
    Dark-subtracted PbI$_2$ images &
    Disorder parameters with ordered structure fixed &
    Diffuse intensity along specular (Bragg rods) and between Bragg peaks along $Q_z$. \\
    \bottomrule
  \end{tabularx}
\end{table}

\begin{enumerate}
  \item \textbf{Beam characterization (direct beam).}
  Direct-beam images are acquired at several detector distances, and the spot centroid and width
  are fit as functions of distance to determine the incident-beam phase-space parameters used in
  the diffraction model (Sec.~\ref{sec:diffraction_model}).

  \item \textbf{Detector geometry calibration (3D powder standard).}
  A polycrystalline hBN calibrant is measured after subtraction of the detector dark frame.
  Multiple Debye--Scherrer rings are fit in detector coordinates to refine detector distance,
  tilts, in-plane rotation, and beam center so that the calibrated rings forward-project to
  concentric circles.

  \item \textbf{Sample alignment and goniometer misalignment.}
  For each sample image, the alignment parameters are determined from indexed Bragg-peak positions
  rather than from global image intensities, because the geometry is constrained primarily by
  feature location. With detector parameters fixed, sample alignment and goniometer misalignment
  are refined by minimizing least-squares residuals between measured and simulated peak positions
  spanning both specular and off-specular regions. Finite-footprint and absorption weighting are
  included through the film dimensions and thickness (Sec.~\ref{sec:diffraction_model}).

  \item \textbf{Mosaicity from Bragg-peak shapes.}
  Two images acquired at different $\theta_i$ are fit simultaneously using the shared mosaic
  parameters of the misorientation model in Sec.~\ref{sec:mosaicity}. The fit is carried out over
  local regions of interest around the indexed reflections, with nuisance scale and background
  terms treated separately in each region so that the shared mosaic kernel is determined by line
  shape rather than by absolute counts.
  Report Lorentzian FWHM $=2\Gamma$, Gaussian FWHM $=2\sqrt{2\ln 2}\,\sigma$, and mixing factor $\eta$.

  \item \textbf{Ordered-structure refinement.}
  With geometry and mosaicity fixed, the ordered-film images of Bi$_2$Se$_3$ and Bi$_2$Te$_3$ are
  used to refine an overall scale together with the structural parameters entering the
  structure-factor intensity model (Eq.~\eqref{eq:structure_factor}), including occupancies,
  anisotropic Debye--Waller terms, and allowed static displacements. The fit is performed against
  the full dark-subtracted 2D images, including the specular region. In a full end-to-end
  implementation, this stage is most naturally formulated as a global intensity refinement with one
  scale per image and physically constrained structural degrees of freedom.

  \item \textbf{Disorder refinement.}
  After the ordered model converges, disorder parameters are introduced and refined against the
  dark-subtracted PbI$_2$ images while the ordered structure is held fixed. This final stage is
  constrained primarily by the diffuse intensity along the specular Bragg rods and between Bragg
  peaks along $Q_z$, and it is kept separate from the ordered refinement to reduce degeneracy
  between average structure and disorder.
\end{enumerate}

All detector images are dark-subtracted before fitting. If additional preprocessing steps such as
flat-field, polarization, solid-angle, or detector-efficiency corrections are applied, they should
be used consistently throughout the workflow; if they are omitted, that choice should be stated
explicitly. Parameter uncertainties are estimated stage by stage by repeated constrained refits
rather than by a single local covariance matrix: calibrant-point selections are resampled for
detector geometry, indexed peak lists are resampled for sample alignment, reflection ROIs are
resampled for mosaicity, and the ordered/disorder intensity fits are repeated from perturbed
initial values. Reported uncertainties are taken as the standard deviation of the accepted
solutions and are propagated forward only after the preceding stage has stabilized.
